---
title: "Extract Layers"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: show
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
```

For Christopher Costello <costello@bren.ucsb.edu>, facilitated by Reniel Cabral <rcabral@bren.ucsb.edu>.

## Install `bbnj` R package

Install the `bbnj` R package (and `devtools` if needed).

```{r, eval=F}
install.packages(devtools)
devtools::install_github("ecoquants/bbnj")
```

## Get Data

See list of datasets in [bbnj reference](http://ecoquants.com/bbnj/reference/index.html#section-datasets). Datasets are "lazy-loaded"" with `library(bbnj)`, meaning available in path but not showing up in Environment tab of RStudio.

```{r}
library(tidyverse)
library(raster)
library(bbnj)

# list bbnj datasets in RStudio:
data(package="bbnj")

# list raster layers within stacks (s_*)
names(s_fish_gfw)
names(s_bio_gmbi)

# get rasters from stacks
profit_fishing <- raster(s_fish_gfw, "mean_scaled_profits_with_subsidies") %>% 
  gap_fill_raster()
nspp_tunas     <- raster(s_bio_gmbi, "nspp_tunas.n.billfishes")
nspp_sharks    <- raster(s_bio_gmbi, "nspp_sharks")

# stack wanted rasters for later cropping
s <- stack(profit_fishing, nspp_tunas, nspp_sharks)
names(s) <- c("profit_fishing", "nspp_tunas", "nspp_sharks")
```

## Get Area of Interest

You'll need to run the `raster::drawExtent()` function interactively (ie not "knitting") within RStudio, outputting to console.

```{r, eval=T}
aoi_rds   <-   "aoi.rds"
s_aoi_rds <- "s_aoi.rds"

if (interactive()){
  # draw your own area of interest
  plot(nspp_tunas)
  aoi <- raster::drawExtent()
  saveRDS(aoi, file = aoi_rds)
} else {
  # load existing
  aoi <- readRDS(aoi_rds) 
}

# show area of interest
aoi

plot(nspp_tunas)
plot(aoi, col = "red", lwd = 2, add = T)

# crop raster
s_aoi <- raster::crop(s, aoi)

# save and plot
saveRDS(s_aoi, file = s_aoi_rds)
plot(s_aoi)
```

