---
title: "Prep climate UBC data"
author: "Ben Best"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = F,
  message = F,
  cache = F
)
```

## Overview

From: Lam, Vicky <v.lam@oceans.ubc.ca> 2019-01-22

1. `Current_MCP1.tif`: Maximum Catch Potential  (MCP) of more than 1,000 marine
fish and invertebrates species in the high seas at the current status (20
years average from 1995 to 2014) under GFDL 8.5 scenario
    - under climate scenario?
    - units in metric tons?
    - list of species?
    - individual species rasters to calculate profitability based on current market prices?

1. `MCP2050_RCP85.tif`: The projected Maximum Catch Potential of more than 1,000
marine fish and invertebrates species in the high seas in the future (20 years
average from 2041to 2060) under GFDL 8.5 scenario
    - Why take average 2041to 2060 vs just 2050? A: to be similar to Current

1. `hsmcpcap851.tif`: The change in MCP in the 2050s relative to the current
status under GFDL 8.5 scenario (I cap the change greater than 2 at 2).
    - Why cap at change > 2?

## Summary Rasters

```{r}
library(tidyverse)
library(raster)
library(fs)
library(glue)
library(here)

par(mar=c(0.1,0.1,0.1,0.1))

dir_ubc <- "~/Gdrive Ecoquants/projects/bbnj/data/raw/UBC-exploited-fish-projections"

tifs <- list.files(dir_ubc, "*.tif$", full.names = T)

cat(paste(basename(tifs), collapse="\n"))
```

### Now & Future

```{r}
library(RColorBrewer)
pal <- colorRampPalette(brewer.pal(11, "Spectral"))
cols <- rev(pal(255))

r_now    <- raster(file.path(dir_ubc, "Current_MCP1.tif"))
r_future <- raster(file.path(dir_ubc, "MCP2050_RCP85.tif"))

r_now
r_future

plot(r_now, col = cols, main="now")
plot(r_future, col = cols, main="future")

hist(r_now, main="now")
hist(r_future, main="future")
```

#### Log Now & Future

```{r}
log(r_now)
log(r_future)

plot(log(r_now), col = cols, main="log(now)")
plot(log(r_future), col = cols, main="log(future)")

hist(log(r_now), main="log(now)")
hist(log(r_future), main="log(future)")
```

### Difference

```{r}
r_dif            <- r_future - r_now
r_dif_clamped0   <- raster(file.path(dir_ubc, "hsmcpcap851.tif"))
r_dif_clamped1   <- r_dif
r_dif_clamped1[r_dif >  2] <-  2
r_dif_clamped1[r_dif < -1] <- -1

r_dif
r_dif_clamped0
r_dif_clamped1

plot(r_dif, col = cols, main="dif")
plot(r_dif_clamped0, col = cols, main="dif_clamped0")
plot(r_dif_clamped1, col = cols, main="dif_clamped1")

hist(r_dif, main="dif")
hist(r_dif_clamped0, main="dif_clamped0")
hist(r_dif_clamped1, main="dif_clamped1")
```

#### Log Difference

```{r}
log(r_dif)

plot(log(r_dif), col = cols, main="log(now)")

hist(log(r_dif), main="log(dif)")
```

## Individual Fish

```{r}
csvs <- list.files(dir_ubc, "*.csv$", recursive = T, full.names = T)
cat(paste(basename(csvs), collapse="\n"))
```

