---
title: "R Notebook"
author: "Ben Best"
date: "1/15/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r load, eval=F}
# libraries ----
library(tidyverse)
library(here)
library(glue)
library(fs)
library(raster)
library(leaflet)
library(sdmpredictors)
select = dplyr::select

# paths ----
# original: ~/github/sdg14/technical/aquamaps_ingest.Rmd|aquamaps_groups.Rmd
dir_am_in        <- "/Volumes/Best HD/mbon_data_big/aquamaps"
spp_csv          <- file.path(dir_am_in, "speciesoccursum_ver0816c.csv")
spp_cells_csv    <- file.path(dir_am_in, "hcaf_species_native_ver0816c.csv")
env_csv          <- file.path(dir_am_in, "hcaf_v6.csv")

dir_ex           <- here("inst/scripts/aquamaps_cc")
spp_examples_csv <- file.path(dir_ex, "spp_examples.csv")
spp_ex_csv       <- file.path(dir_ex, "spp_ex.csv")
spp_ex_cells_csv <- file.path(dir_ex, "spp_ex_cells.csv")

env_lyrs_csv     <- file.path(dir_ex, "env_lyrs.csv")

# create half-deg raster ----
r_na <- raster(
  xmn = -180, xmx = 180, ymn = -90, ymx = 90, 
  resolution=0.5, crs=leaflet:::epsg4326)
r_a <- area(r_na)

# read data ----
if (any(!file.exists(c(spp_examples_csv, spp_ex_csv, spp_ex_cells_csv)))){
  spp_examples <- tribble(
    ~common           , ~speciesid,
    "right whale"     , "ITS-Mam-180537",
    "blue whale"      , "ITS-Mam-180528",
    "cusk eel"        , "Fis-50764",
    "bluefin trevally", "Fis-29762",
    "skipjack tuna"   , "Fis-22777",
    "blue shark"      , "Fis-23222")
  write_csv(spp_examples, spp_examples_csv)
  
  read_csv(spp_csv) %>% 
    filter(SPECIESID %in% spp_examples$speciesid) %>% 
    write_csv(spp_ex_csv)
    
  read_csv(spp_cells_csv) %>% 
    filter(SpeciesID %in% spp_examples$speciesid) %>% 
    write_csv(spp_ex_cells_csv)
}
spp_examples <- read_csv(spp_examples_csv)
spp          <- read_csv(spp_ex_csv)
spp_cells    <- read_csv(spp_ex_cells_csv)
env          <- read_delim(env_csv, delim = "\t")
```

# AquaMaps Environmental

[AquaMaps Environmental Dataset](https://www.aquamaps.org/main/envt_data.php)

```{r lyrs, eval=F}
list_layers() %>% 
  write_csv(file.path(dir_ex, "sdm_lyrs.csv"))
list_layers_future() %>% 
  write_csv(file.path(dir_ex, "sdm_lyrs_future.csv"))
sdmpredictors_lyrs %>% as_tibble()

tibble(
  am_layer = names(env)) %>% 
    write_csv(env_lyrs_csv)
```


`Bio-ORACLE`, `model "AOGCM" scenario "RCP85" year 2050, version 2`:
		
- Depth	( Mean, Minimum, Maximum )
  - Depth Mean (m): [0, 10000] - `BO_bathymean`, ``
  - Depth Minimum (m): [0, 10000] - `BO_bathymin`, ``
  - Depth Maximum (m): [0, 10000] - `BO_bathymax`, ``
- Temperature	( Surface: Mean, Minimum, Maximum; Bottom: Mean )
  - Temperature: Surface Mean (°C): [-2, 33] - `BO2_tempmean_bdmin`, `BO2_RCP85_2050_tempmean_bdmean`
  - Temperature: Surface Minimum (°C): [-2, 33] - `BO2_tempmin_bdmin`, ``
  - Temperature: Surface Maximum (°C): [-2, 33] - `BO2_tempmax_bdmin`, ``
  - Temperature: Bottom Mean (°C): [-2, 33] - `BO2_tempmean_bdmax`, ``
- Salinity	( Surface: Mean, Minimum, Maximum; Bottom: Mean )
  - Salinity: Surface Mean (psu): [0, 41] - ``, ``
  - Salinity: Surface Minimum (psu): [0, 41] - ``, ``
  - Salinity: Surface Maximum (psu): [0, 41] - ``, ``
  - Salinity: Bottom Mean (psu): [0, 41] - ``, ``
- Primary Productivity (Mean)
  - Primary Productivity: Mean (mgC·m-²·day -1): [0, 6000] - ``, ``
- Sea Ice Concentration (Mean)
  - Sea Ice Concentration: Mean: [0.01, 1] - ``, ``
  
  
```{r, eval=F}
env_lyrs <- read_csv(env_lyrs_csv)
env_lyrs <- bind_rows(
  env_lyrs %>% 
    filter(!has_future),
  env_lyrs %>% 
    filter(has_future) %>% 
    mutate(
      bo_rcp85_2050 = get_future_layers(bo_layer, scenario = "RCP85", year = 2050) %>% pull(layer_code),
      bo_rcp85_2100 = get_future_layers(bo_layer, scenario = "RCP85", year = 2050) %>% pull(layer_code)))
write_csv(env_lyrs, env_lyrs_csv)
```

## Example Species

- [Abyssobrotula galatheae](https://www.aquamaps.org/ShowMapParam.php?SpecID=Fis-50764&user_session=58): deepest-living fish known in abyssal plain
Water temp. (°C) (bottom): -0.955, 1.06, 2.57, 7.35 (Min_A, Min_P, Max_P, Max_A)

//input

```{r}
library(rvest)


spid <- "Fis-50764"

spp_cc_examples

get_sp_params <- function(spid){
  # TODO: Area restrictions: FAO Area(s): 21, 27, 31, 34, 47, 51, 57, 61, 71, 77, 81, 87; Pelagic: True; Bounding Box (NSWE):
  # spid="Fis-50764"
  
  tmp <- tempfile()
  
  # get first expert listing
  url <- glue("https://www.aquamaps.org/preMap2.php?SpecID={spid}")
  download.file(url, destfile = tmp, quiet=TRUE)
  html <- read_html(tmp)
  url <- html %>% 
    html_node("td a[id=link]") %>% 
    html_attr("href")
  url <- glue("https:{url}")
  
  # get habitat params url
  download.file(url, destfile = tmp, quiet=TRUE)
  url <- read_html(tmp) %>% 
    html_nodes("form input[name=legend]") %>% 
    html_attr("value") %>% 
    str_split("\\n", simplify=T) %>% 
    str_subset("mapparam_window = window.open") %>% 
    str_replace('.*mapparam_window = window.open\\("(.*)","(.*)","(.*)"\\);', "\\1") %>% 
    str_trim()
  url <- glue("https:{url}")

  # read habitat params
  download.file(url, destfile = tmp, quiet=TRUE)
  nodes <- read_html(tmp) %>% 
    html_nodes("table td input")
  suffixes <- c("YN", "Min","PrefMin","PrefMax","Max")
  regex_sfx <- paste0(paste(suffixes, collapse="$|"), "$")
  
  params <- read_html(tmp) %>% 
    html_nodes(xpath='//td[input[@type="checkbox"]]/following-sibling::td[1]') %>% 
    html_text(trim=T) %>% 
    str_replace_all("\\r|\\n|\\t", "") %>% # works
    str_replace_all("  ", " ") # works

  d <- tibble(
    name  = html_attr(nodes, "name"),
    type  = html_attr(nodes, "type"),
    value = html_attr(nodes, "value") %>% as.numeric()) %>% 
    filter(!type %in% c("hidden")) %>% 
    mutate(
      name = str_replace_all(name, "_edit$", ""),
      sfx  = str_extract(name, regex_sfx),
      var  = map2_chr(name, sfx, ~ str_replace(.x, .y, ""))) %>% 
    select(-name, -type)
  
  var_ordered <- unique(d$var)
  
  d <- d %>% 
    spread(sfx, value)
  
  d <- d %>% 
    mutate(
      param      = !!params[match(d$var, var_ordered)],
      spid       = !!spid,
      url_params = url) %>% 
    select(spid, var, param, one_of(suffixes), url_params)
  
  unlink(tmp)

  d
}

spp_params <- map_df(spp_cc_examples, get_sp_params)

table(spp_params$param)
View(spp_params)

spp_cc_examples
unique(spp_params$url_params)
```

- [Caranx melampygus](https://www.aquamaps.org/ShowMapParam.php?SpecID=Fis-29762&user_session=87) (bluefin trevally): example of pelagic fish
Water temp. (°C) (surface): 21.5, 26, 28.95, 33.15 (Min_A, Min_P, Max_P, Max_A)

- [Sea Around Us | Fisheries, Ecosystems and Biodiversity](http://www.seaaroundus.org/data/#/eez/971?chart=catch-chart&dimension=taxon&measure=tonnage&limit=10)

- [right whale (_Eubalaena glacialis_)](https://www.aquamaps.org/preMap2.php?cache=1&SpecID=ITS-Mam-180537): expert contracted range

- [blue whale (_Balaenoptera musculus_)](https://www.aquamaps.org/preMap2.php?cache=1&SpecID=ITS-Mam-180528): global range

- [cusk eel (_Abyssobrotula galatheae_)](https://www.aquamaps.org/ShowMapParam.php?SpecID=Fis-50764&user_session=58): deepest-living fish known in abyssal plain

- [bluefin trevally (_Caranx melampygus_)](https://www.aquamaps.org/ShowMapParam.php?SpecID=Fis-29762&user_session=87) (): example of pelagic fish
  
- [skipjack tuna (_Katsuwonus pelamis_)](https://www.aquamaps.org/preMap2.php?cache=1&SpecID=Fis-22777)

- [blue shark (_Prionace glauca_)](https://www.aquamaps.org/preMap2.php?cache=1&SpecID=Fis-23222)



```{r}
spp_examples
head(spp)
head(env)
head(spp_cells)




spp <- spp %>% 
  filter(SPECIESID %in% spp_cc_examples)
spp_cells <- spp_cells %>% 
  filter(SPECIESID %in% spp_cc_examples)

write_rds()
```


