---
title: "Taxonomic Group Treemap"
author:
date: "7/31/2019"
output:
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, message=F)
```
####Taxonomic groups from [gmbi package](https://github.com/marinebon/gmbi/blob/master/vignettes/calc.Rmd) used in [bbnj analysis](https://github.com/ecoquants/bbnj/)
```{r libraries}
suppressPackageStartupMessages({
  library(dplyr)
  library(treemap)
  library(ggplot2)
  library(kableExtra)
})
```

``` {r functions}

## filter spp dataset by groups number
filter_group <- function (df, groupnumber) {
    df %>%
    filter(!is.na(UQ(sym(groupnumber))))
}

## summarise by taxonomic group name

summarise_group <- function(df, groupnumber) {
  groupnumber<- as.symbol(groupnumber)
  groupnumber<- enquo(groupnumber)
  df %>% 
  group_by((!!groupnumber)) %>%
  summarize(n = n())
}

```

``` {r}
# Read in spp_bbnj.csv with groupings:
  # groups00: All original AquaMaps species 
  # groups01: 1st taxonomic grouping [n = 23]
  # groups02: 1st grouping, species (prob≥ 0.5) in high seas
  # groups03: 1st grouping, species (prob ≥ 0.5) and 25% of its range in high seas
  # groups04: 2nd taxonomic grouping [n = 9]
  # groups05: 2nd grouping, species (prob≥ 0.5) in high seas
  # groups06: 2nd grouping, species (prob ≥ 0.5) and 25% of its range in high seas

spp <- readr::read_csv("extdata/spp_bbnj.csv") 
spp[spp=="Non-squid cephalopods"]<-"Cephalopods"

```

```{r group 01}

groupnumber<-"groups01"

spp_grp<-filter_group(spp,groupnumber)
nspp_grp<-nrow(spp_grp)
taxa_grp <-summarise_group(spp_grp,groupnumber)
ngrp_grp <-nrow(taxa_grp)
```
### Groups v01
____
- All AquaMaps (no HS filter)

- `r ngrp_grp` Taxonomic Groups

- `r nspp_grp` Species

```{r out.width = '1200px'}

#Create treemap where size of rectangle corresponds to number of species in each group

taxa_treemap<- treemap(taxa_grp,
                       index=groupnumber,
                       vSize="n",
                       type="index",
                       force.print.labels = TRUE,
                       title="Number of species per group"
)

#Create table of nspp for each taxonomic group
knitr::kable(taxa_grp,
             col.names = c("Group",
                           "Number of Species")) %>% 
  kable_styling(bootstrap_options = "striped")
```


```{r group 02}

groupnumber<-"groups02"

spp_grp<-filter_group(spp,groupnumber)
nspp_grp<-nrow(spp_grp)
taxa_grp <-summarise_group(spp_grp,groupnumber)
ngrp_grp <-nrow(taxa_grp)
```
### Groups v02
____
- AquaMaps species (prob≥ 0.5) in high seas

- `r ngrp_grp` Taxonomic Groups

- `r nspp_grp` Species

```{r out.width = '1200px'}

#Create treemap where size of rectangle corresponds to number of species in each group

taxa_treemap<- treemap(taxa_grp,
                       index=groupnumber,
                       vSize="n",
                       type="index",
                       force.print.labels = TRUE,
                       title="Number of species per group"
)

#Create table of nspp for each taxonomic group
knitr::kable(taxa_grp,
             col.names = c("Group",
                           "Number of Species")) %>% 
  kable_styling(bootstrap_options = "striped")
```


```{r group 03}

groupnumber<-"groups03"

spp_grp<-filter_group(spp,groupnumber)
nspp_grp<-nrow(spp_grp)
taxa_grp <-summarise_group(spp_grp,groupnumber)
ngrp_grp <-nrow(taxa_grp)
```
### Groups v03
____
- AquaMaps species (prob≥ 0.5) in high seas and 25% of its range in high seas

- `r ngrp_grp` Taxonomic Groups

- `r nspp_grp` Species

```{r out.width = '1200px'}

#Create treemap where size of rectangle corresponds to number of species in each group

taxa_treemap<- treemap(taxa_grp,
                       index=groupnumber,
                       vSize="n",
                       type="index",
                       force.print.labels = TRUE,
                       title="Number of species per group"
)

#Create table of nspp for each taxonomic group
knitr::kable(taxa_grp,
             col.names = c("Group",
                           "Number of Species")) %>% 
  kable_styling(bootstrap_options = "striped")
```


```{r group 04}

groupnumber<-"groups04"

spp_grp<-filter_group(spp,groupnumber)
nspp_grp<-nrow(spp_grp)
taxa_grp <-summarise_group(spp_grp,groupnumber)
ngrp_grp <-nrow(taxa_grp)
```
### Groups v04
____
- All AquaMaps (no HS filter)

- `r ngrp_grp` Taxonomic Groups

- `r nspp_grp` Species

```{r out.width = '1200px'}

#Create treemap where size of rectangle corresponds to number of species in each group

taxa_treemap<- treemap(taxa_grp,
                       index=groupnumber,
                       vSize="n",
                       type="index",
                       force.print.labels = TRUE,
                       title="Number of species per group"
)

#Create table of nspp for each taxonomic group
knitr::kable(taxa_grp,
             col.names = c("Group",
                           "Number of Species")) %>% 
  kable_styling(bootstrap_options = "striped")
```


```{r group 05}

groupnumber<-"groups05"

spp_grp<-filter_group(spp,groupnumber)
nspp_grp<-nrow(spp_grp)
taxa_grp <-summarise_group(spp_grp,groupnumber)
ngrp_grp <-nrow(taxa_grp)
```
### Groups v05
____
- AquaMaps species (prob≥ 0.5) in high seas

- `r ngrp_grp` Taxonomic Groups

- `r nspp_grp` Species

```{r out.width = '1200px'}

#Create treemap where size of rectangle corresponds to number of species in each group

taxa_treemap<- treemap(taxa_grp,
                       index=groupnumber,
                       vSize="n",
                       type="index",
                       force.print.labels = TRUE,
                       title="Number of species per group"
)

#Create table of nspp for each taxonomic group
knitr::kable(taxa_grp,
             col.names = c("Group",
                           "Number of Species")) %>% 
  kable_styling(bootstrap_options = "striped")
```


```{r group 06}

groupnumber<-"groups06"

spp_grp<-filter_group(spp,groupnumber)
nspp_grp<-nrow(spp_grp)
taxa_grp <-summarise_group(spp_grp,groupnumber)
ngrp_grp <-nrow(taxa_grp)
```
### Groups v06
____
- AquaMaps species (prob≥ 0.5) in high seas and 25% of its range in high seas

- `r ngrp_grp` Taxonomic Groups

- `r nspp_grp` Species

```{r out.width = '1200px'}

#Create treemap where size of rectangle corresponds to number of species in each group

taxa_treemap<- treemap(taxa_grp,
                       index=groupnumber,
                       vSize="n",
                       type="index",
                       force.print.labels = TRUE,
                       title="Number of species per group"
)

#Create table of nspp for each taxonomic group
knitr::kable(taxa_grp,
             col.names = c("Group",
                           "Number of Species")) %>% 
  kable_styling(bootstrap_options = "striped")
```


